---
title: "Assignment 4 Task 2"
author: "Nicholas Bissonnette"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE , warning= FALSE)
library(tidyverse)
library(broom)
library(lubridate)
library(here)
library(beeswarm)
library(ggbeeswarm)
library(effsize)
library(kableExtra)
```

## 1.Introduction


For this assignment I will look at data collected from the Bonanza Creek experimental Forest. The forest is located in Fair banks, Alaska and they conduct various forms of research from ecosystem dynamics to effects of climate change in a boreal biome. Bonanza Creek is part of the long term ecological research network which is supported by the national science foundation  
data on Snowshoe Hares was collected from 1998-2012. I will particularly look at juvenile hares in the study area. 


## 2.Data and analysis


First I will look at study design and the distribution of collection over time. Then I will look at the difference in weights among the sexes as they relate to the trapping sites. Lastly I will look at the linear relationship between weight and hind foot length of juvenile snowshoe hares. The statiscal tools used are t-tests, effect size, liner models, and regression.     

**CITATION**

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14 (Accessed 2020-11-17).



## 3.Exploratory findings


```{r, message=FALSE}
#Here  I will read in the dara using the here() and read_csv()

hares <- read_csv(here("data","bonanza_hares.csv"))

# Here is where I wrangle data to make different data frames, one for years, and one for age
hares_year <- hares %>% 
  mutate(Date = mdy(date)) %>% 
  mutate(year = year(Date))

hares_j <- hares_year %>% 
  filter(age =="j") %>% 
  group_by(year)

hares_j_count <- hares_year %>% 
  filter(age =="j") %>% 
  count(age, year)
# this where I make a summary table of mean traps and median traps

hares_sum <- hares_j_count %>% 
  summarise(mean_trap = mean( n),
            median_trap= median(n))

```

**A.** 

```{r, message=FALSE}
#This is a histogram to explor distribution

ggplot(data=hares_j, aes(x=year))+
  geom_histogram ( aes(fill= "red", na.rm = TRUE))+
  labs(title= "Annual juvenile hare trap counts by site ", x= "Year" , y= "# of trapped juvenile hares" ,
       subtitle = "Figure 1.1")+
  theme_minimal()+
  theme(axis.text.x= element_text(angle= 90, hjust=1))+
  theme(legend.position = "none")

```


### **Figure 1.1**  shows the counts of each trapped juvenile  hare at the  three trapping sites ( bonbos, bonmat, bonrip) by year.




The data tells us that a total of 378 hares were trapped from 1999 -2012. The year with the lowest amount traps was 2010 with 2 and the year with the most traps was 1999 with 126. There are also two years with no data. This data appears to not be normally distributed, with the majority of traps coming in the early years of the study. The mean annual count of traps is 31.5, which is not close to a value that is seen in any individual year.The median value of the annual traps is 18.5. This large difference between median and mode is evidence for none normal distribution.     




**B.**


Next, I will create a data visualization to compare juvenile hare weights by sex and site. I will do this making a dataframe of sex, age, weight and grid. Then I will plot that data in a facet version to show each.   

```{r, message= FALSE}
# here is where I wrangle data to create a data frame of the desired categories (select()) and filter() to get only juveniles included. 
hare_sex <- hares %>% 
  select(age, grid, weight, sex) %>% 
  filter(age == "j")

#using the above dataframe I created a summary table by grouping sex and site using group_by() and then summarize() to get mean, sd and sample size of grouped data. 
weight_sum <-hare_sex %>% 
  group_by(sex, grid) %>% 
  summarize(
    mean_weight= mean(weight, na.rm=T),
    sd_weight =sd(weight, na.rm=T),
    sample_size= n())

```

```{r}
# I chose beeswarm plat here to better see individual points and I put the mean values on each plot woth error pars to get a better ideal of central tendency between the sites. Using facet_wrap() helps show each site individually. 

ggplot() +
  geom_beeswarm(data=hare_sex, 
                aes (x=sex, y=weight),
                color= "blue",
                size=1,
                corral=c("wrap")
                )+
  facet_wrap(~grid)+
  labs(title = "Snowshoe hare weight and sex compaison by trap site", subtitle = "Figure 1.2", y= "Weight in grams", x= "Hare Sex")+
  geom_point(data= weight_sum, aes(x=sex, y= mean_weight),
             size=4) +
  geom_errorbar(data= weight_sum,
                aes(x=sex,
                    ymax=mean_weight +sd_weight,
                    ymin= mean_weight - sd_weight),
                width=1)
   

```



### **Figure 1.2** This shows the weight(grams) compared to hare sex at each of the trap sites. 

```{r, warning= FALSE, fig.show= 'hide'}
# I made this quantile plot to see normal distribution between the sites. 
ggplot(data= hare_sex, aes(sample= weight)) +
  geom_qq(aes(color=grid))+
  facet_wrap(~grid, scales= "free")+
  theme(legend.position = "none")
```

Figure 1.2 shows that the majority of data is coming from the bonrip site. The bonbs sate has the smallest sample size. A quantile-quantile plot shows a linear relationship  for the bonrip trap site, bonmat site shows less linear relationship and the bonbs shows the least linear relationship. The more linear a relationship is the more normally distributed the data is which suggests that bonrip is the most normal and bonbs is the least normal.  




**C.**


Here I will make a summary statistic table for the hares of the weights. This information can be used to find mean, standard deviation, and sample size of the snowshoe hares.

### **Table 1.1** The below table show the mean, Standard Deviation, and Sample Size of Snowshoe Hare weights broken up by sex.

```{r, message= FALSE}
#this is where I made a tabe of summary states like before but I also wanted to included this in the knitted doc os I used kbl() for show and used add_header_above() to add a title.
weight_sum <-hare_sex %>% 
  group_by(sex) %>% 
  summarize(
    mean_weight= mean(weight, na.rm=T),
    sd_weight =sd(weight, na.rm=T),
    sample_size= n())
weight_sum %>% 
  kbl() %>% 
  kable_paper() %>% 
  add_header_above(c("Mean, Standard Deviation, and Sample Size of Snowshoe Hare weights "=4))
```






```{r, message=FALSE, results= 'hide'}
#this is where I vade two vecotrs to preform the t-test and cohen's d
hares_f <- hares_j %>% 
  filter(sex== "f") %>% 
  pull(weight)

hares_m <- hares_j %>% 
  filter(sex =="m") %>% 
  pull(weight)

hares_ttest <- t.test(hares_f, hares_m)


cohen.d(hares_f, hares_m, na.rm = T)
```

**D.**


After running the Two Sample T-test, I can see that the difference in mean weight between female and male snowshoe hares is 90.47. The Cohen's d test shows a value 0.29 which is a low moderate effect, indicating that there is a difference and it would take effort to study what that difference is. The p-value fro the the t-test is 0.007 which show a relatively low chance that samples came from populations with the same mean. 



**E.**


Here I will explore the relationship between juvenile hare weight and hind foot length. 

```{r, message=FALSE}
# this is where I made a scatter plot to explore linear regression. Using the smooth(method="lm") add the regression line on the plot. 
ggplot(data=hares_j, aes(x=hindft, y= weight))+
  geom_point(stat="identity")+
  labs(title = "Weight and hind foot length comparison of Juvenile Snowshoe Hares",subtitle = "Figure 1.3", x="Hind foot length (mm)", y= "Weight (g)")+
  geom_smooth(method = "lm", se=FALSE)
  
```


### **Figure 1.3** This figure shows the relationship between hind foot length (mm) and weight (g) of Juvenile Snowshoe Hares across all trap locations. 




```{r}
#this is where I made the linear model
hares_lm <- lm(weight ~ hindft, data = hares_j)



```

The linear model for the data in figure 1.4 shows that the slope is `r round(hares_lm$coefficient[2],2)` (g / mm)
and the y-intercept is `r round(hares_lm$coefficient[1],2)` (g). The full equation is weight = `r round(hares_lm$coefficient[2],2)`*(hind foot length) + (`r round(hares_lm$coefficient[1],2)`)


```{r}
#this is where the person's R test is preformed 
hares_cor <-cor.test(hares_j$weight, hares_j$hindft)
```

Pearson's R will give us the correlation for figure 1.3. I am able to see that the Pearson's r correlations is `r round(hares_cor$estimate,2)`. This value indicates a moderate positive correlation. I would say that the higher the values are for hind foot length and weight, the more the correlation decreases. This indicates that there is a week relationship between weight and hind foot length as the overall size increases. 


## 4.Summary

After exploring the data of Juvenile Snowshoe Hares, I have been able to observe differences in males and female weights as they relate to site trappings. Similar mean weight values were observed at all three trapping locations. The difference between means in both male and female populations is significant but the difference is low-moderate as shown by the effect size. The  linear model shows a moderate positive correlation, but the scatter plot shows a higher amount of non-correlation the further higher in value the data gets.  